Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-11-03 manuscript/manuscript.qmd

set.seed(888)
```

@@

```{r fig-vocabulary}
#| label: fig-vocabulary
#| fig-cap: "Participant vocabulary sizes across ages and language profiles. Vocabulary size scores are presented using two measures: L1 vocabulary size (proportion of words marked as *Understands* in the vocabulary checklist of the dominant language), and total vocabulary size (proportion of words marked as *Understands* in vocabulary checklists of both dominant and non-dominant languages). Grey lines connect vocabulary size scores from the recurrent participants. For visualisation purposes, participants were group into categorical age groups at 21, 25, 30 months, the ages around which most participants were tested."
#| fig-width: 8
#| fig-height: 4.5
vocabulary_bcn |> 
	pivot_longer(c(total_prop, l1_prop),
				 names_to = "measure",
				 values_to = "prop",
				 names_transform = \(x) gsub("_prop", "", x)) |> 
	mutate(measure = if_else(measure=="l1", 
							 "L1 vocabulary", 
							 "Total vocabulary"),
		   is_imputed = if_else(is_imputed, "Imputed", "Observed") |> 
		   	factor(levels = c("Observed", "Imputed"))) |> 
	ggplot(aes(age_group, prop,
			   colour = lp,
			   fill = lp)) +
	facet_wrap(~measure) +
	geom_line(aes(group = child_id),
			  colour = "grey",
			  alpha = 1/4, linewidth = 3/4) +
	geom_dots(aes(side = if_else(lp=="Monolingual", "left", "right")),
			  dotsize = 2,
			  scale = 0.5,
			  alpha = 3/4) +
	stat_summary(fun.data = mean_se,
				 geom = "errorbar",
				 linewidth = 3/4,
				 width = 0.15,
				 colour = "black",
				 show.legend = FALSE,
				 position = position_dodge(width = 1.15)) +
	stat_summary(fun = mean,
				 geom = "point",
				 size = 2,
				 colour = "black",
				 show.legend = FALSE,
				 position = position_dodge(width = 1.15)) +
	# geom_point(alpha = 1) +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(labels = scales::percent) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  legend.title = element_blank())

@@

>>>>

set.seed(1234)
```

@@

```{r fig-vocabulary}
#| label: fig-vocabulary
#| fig-cap: "Participant vocabulary sizes across ages and language profiles. Vocabulary size scores are presented using two measures: L1 vocabulary size (proportion of words marked as *Understands* in the vocabulary checklist of the dominant language), and total vocabulary size (proportion of words marked as *Understands* in vocabulary checklists of both dominant and non-dominant languages). Grey lines connect vocabulary size scores from the recurrent participants. For visualisation purposes, participants were group into categorical age groups at 21, 25, 30 months, the ages around which most participants were tested."
#| fig-width: 8
#| fig-height: 4.5
vocabulary_bcn |> 
	pivot_longer(c(total_prop, l1_prop),
				 names_to = "measure",
				 values_to = "prop",
				 names_transform = \(x) gsub("_prop", "", x)) |> 
	mutate(measure = if_else(measure=="l1", 
							 "L1 vocabulary", 
							 "Total vocabulary"),
		   is_imputed = if_else(is_imputed, "Imputed", "Observed") |> 
		   	factor(levels = c("Observed", "Imputed"))) |> 
	ggplot(aes(age_group, prop,
			   colour = lp,
			   fill = lp)) +
	facet_wrap(~measure) +
	geom_line(aes(group = child_id),
			  colour = "grey",
			  alpha = 1/4, linewidth = 3/4) +
	geom_dots(aes(side = if_else(lp=="Monolingual", "left", "right")),
			  dotsize = 2,
			  scale = 0.5,
			  alpha = 3/4) +
	stat_summary(fun.data = mean_se,
				 geom = "errorbar",
				 linewidth = 3/4,
				 width = 0.15,
				 colour = "black",
				 show.legend = FALSE,
				 position = position_dodge(width = 1.15)) +
	stat_summary(fun = mean,
				 geom = "point",
				 size = 2,
				 colour = "black",
				 show.legend = FALSE,
				 position = position_dodge(width = 1.15)) +
	# geom_point(alpha = 1) +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(labels = scales::percent) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  legend.title = element_blank())
```

@@

# Study 2

```{r}
#| label: participants-numbers-oxf
participants_oxf <- filter(participants, location=="Oxford")

n_participants_total <- nrow(distinct(participants_oxf, child_id))
n_sessions_total <- count(participants_oxf)

n_participants_sex <- participants_oxf |> 
	distinct(sex, child_id) |> 
	count(sex)

n_participants_sessions <- participants_oxf |> 
	count(child_id, name = "n_sessions") |> 
	count(n_sessions) |> 
	group_split(n_sessions) |> 
	set_names(paste0("session_", 1:2)) |> 
	map("n")

ages <- participants_oxf |> 
	summarise(n = n(),
			  n_participants = n_distinct(child_id),
			  across(age, lst(mean, sd, min, max)),
			  across(matches("doe_"), lst(mean, sd)),
			  age_vctr = list(age),
			  .by = c(lp, test_language))

We collected data from `r n_participants_total$n[1]` children (`r n_participants_sex[1]` female, `r n_participants_sex[2]` male, `r n_participants_sex[3]` sex not reported, age: *Mean* = `r round(ages$age_mean, 2)` months, *SD* = `r round(ages$age_sd, 2)`, *Range* = [`r round(ages$age_min, 2)`-`r round(ages$age_max, 2)`]), living in the Oxfordshire area (United Kingdom), tested at the Oxford BabyLab at the University of Oxford. Families were recruited from maternity rooms in private hospitals and social media, and contacted via phone when the child's age spanned between 20 and 32 months. From the `r n_participants_total$n` children that participated, `r n_participants_sessions$session_1` participated once, and `r n_participants_sessions$session_2` participated twice. Recurrent participants were tested with at least XXX months of difference. We gathered a total of `r n_sessions_total$n` testing sessions. All participants were being raised in exclusively British English monolingual homes.

```{r vocab-values-oxf}
vocabulary_oxf <- inner_join(vocabulary, 
							 select(participants_oxf, session_id),
							 by = join_by(session_id))
```

We collected vocabulary data using parental responses to the Oxford Communicative Development Inventory (OCDI) [@hamilton2000infant]. Families were sent the questionnaire immediately after each experimental session, and were given two weeks to fill it. Since all vocabulary sizes of English monolingual infants were only assessed in English, both *L1 vocabulary* and *total vocabulary* sizes were identical (see @fig-vocabulary-oxf).

```{r fig-vocabulary-oxf}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant vocabulary sizes across ages and language profiles. Vocabulary size scores are presented using two measures: L1 vocabulary size (proportion of words marked as *Understands* in the vocabulary checklist of the dominant language), and total vocabulary size (proportion of words marked as *Understands* in vocabulary checklists of both dominant and non-dominant languages). Vocabulary size scores from the recurrent participants are shown linked."
#| eval: false
#| fig-width: 4
#| fig-height: 4
vocabulary_oxf |> 
	inner_join(select(participants, child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |> 
	ggplot(aes(age_group, l1_prop)) +
	geom_line(aes(group = child_id),
			  colour = "grey",
			  alpha = 1/4, linewidth = 3/4) +
	geom_dots(dotsize = 2,
			  colour = clrs[5],
			  fill = clrs[5],
			  scale = 0.5,
			  side = "both",
			  alpha = 3/4) +
	stat_summary(fun.data = mean_se,
				 geom = "errorbar",
				 linewidth = 3/4,
				 width = 0.1,
				 colour = "black",
				 show.legend = FALSE) +
	stat_summary(fun = mean,
				 geom = "point",
				 size = 2,
				 colour = "black",
				 show.legend = FALSE) +
	# geom_point(alpha = 1) +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(labels = scales::percent) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  legend.title = element_blank())
```

New stimuli lists were generated to adapt the experimental procedure to English-learning infants. Stimuli lists were created following the same constraints as in Study 1. Auditory stimuli were recorded following the same procedure as in Study 1, produced by Southern British English native female speaker in a toddler-directed manner. Audio and picture processing were performed following the same workflow as in Study 1.

### Procedure and apparatus

Same as in Study 1. The study was run using a custom Matlab script, PresentMate, based on the PsychToolbox-3 extension [@kleiner2007s; @brainard1997psychophysics; @pelli1997videotoolbox]. Visual fixations were recorded using a Tobii TX 300 eye-tracker and a 23-in screen of 1920 $\times$ 1080 resolution. The Tobii Analytics SDK 3.0 was used to interact with the eye-tracking while the experiment was running. Sampling rate was set at 120 Hz. A 9-point calibration was performed before every experimental session, in which the picture of a colourful beach ball was presented.

Same as in Study 1, but removing the *Group* predictor from the models, as only monolingual infants were tested.

```{r results-oxf}
var_labels <- c("b_Intercept" = "Intercept",
				"b_age_std" = "Age (+1 SD)",
				"b_condition1" = "Condition (U vs. RN)",
				"b_condition1" = "Condition (RN vs. RC)",
				"sd_child_id__Intercept" = "Intercept",
				"sd_child_id__age_std" = "Age",
				"sd_child_id:session_id__Intercept" = "Intercept",
				"sd_child_id:session_id__age_std" = "Age",
				"sd_child_id:session_id__lpBilingual" = "Group")

coefs <- map(model_fits_oxf, 
			 \(x) get_coefs(model = x, var_labels,
			 			   `b_.*`, `sd_.*`, regex = TRUE))

loos_tbl <- loo_compare(model_loos_oxf) |> 
	as.data.frame() |> 
	rownames_to_column("model") |> 
	select(model, matches("diff|elpd"))

```{r tbl-loos-oxf}
loos_tbl |> 
	mutate(model = case_when(model=="fit_oxf_0" ~ "$\\mathcal{M}_{0}$",
							 model=="fit_oxf_1" ~ "$\\mathcal{M}_{1}$",
							 model=="fit_oxf_2" ~ "$\\mathcal{M}_{2}$",
							 model=="fit_oxf_3" ~ "$\\mathcal{M}_{3}$")) |>
	gt() |> 
	fmt_number(is.numeric) |> 
	# fmt_markdown(is.character, ) |> 
	cols_label(elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD")
```

```{r timecourse-oxf}
timebin_values <- make_std(seq(0, 17, length.out = 100),
						   mean(data_oxf$timebin),
						   sd(data_oxf$timebin))

age_values <- make_std(c(20, 30),
					   mean(data_oxf$age),
					   sd(data_oxf$age))

nd <- expand_grid(distinct(data_oxf, condition),
				  timebin_std = timebin_values,
				  age_std = age_values)

epreds <- add_epred_draws(nd, model_fits_oxf$fit_oxf_2, 
						  re_formula = NA,
						  ndraws = NULL,
						  value = ".elog") |> 
	mutate(age_std = factor(age_std,
							levels = age_values, 
							labels = paste0(c(20, 30), " months")))

epreds_chance <- epreds |> 
	median_hdci() |> 
	mutate(is_lower_positive = .lower > 0) |> 
	filter(is_lower_positive) |> 
	summarise(over_chance = min(timebin_std),
			  .by = c(condition, age_std)) |> 
	mutate(over_chance = (over_chance * 
						  	sd(data_oxf$timebin) +
						  	mean(data_oxf$timebin))*100+300)

epreds_chance_report <- epreds_chance |> 
	mutate(over_chance = sprintf("%.2f", over_chance),
		   age_std = paste0("age_", gsub(" months", "", age_std))) |> 
	arrange(age_std) %>%
	split(tolower(.$age_std)) 
```

```{r fig-epreds-oxf}
#| label: fig-epreds-oxf
#| fig-cap: "Posterior mean prdictions of the time course of target fixation in the test phase. Lines and intervals indicate the median and 95% credible interval of the posterior predictions, respectively."
#| fig-width: 8
#| fig-height: 4
epreds |> 
	ggplot(aes(timebin_std, .elog,
			   colour = condition,
			   fill = condition)) +
	facet_wrap(~age_std) +
	geom_hline(yintercept = 0, linetype = "dashed") +
	stat_summary(fun.data = mean_qi,
				 geom = "ribbon",
				 alpha = 1/2,
				 linewidth = NA) +
	stat_summary(fun = mean,
				 geom = "line",
				 linewidth = 3/4) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_x_continuous(breaks = make_std(seq(0, 17, 3),
										 mean(data_bcn$timebin),
										 sd(data_bcn$timebin)),
					   labels = \(x) format(seq(3e2, 2e3, 3e2),
					   					 big.mark = ",")) &
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  legend.position = "top")
```

```{r fig-recognition-oxf}
#| label: fig-recognition-oxf
#| fig-cap: "Model-predicted recognition times for monlingual and bilinguals at 20 and 30 months of age."
#| fig-width: 7
#| fig-height: 3
epreds_chance |> 
	ggplot(aes(age_std, over_chance,
			   colour = sort(condition),
			   fill = sort(condition))) + 
	geom_point(shape = 16, size = 4) +
	geom_line(aes(group = condition),
			  linewidth = 1) +
	# geom_point(size = 3) +
	labs(x = "Group",
		 y = "Recognition time (ms)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_y_continuous(breaks = seq(0, 1e3L, 25)) +
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  panel.grid.minor.y = element_line(colour = "grey",
		  								  linetype = "dotted"),
		  legend.position = "top",
		  axis.title.x = element_blank())
```

## Discussion

# General discussion
